Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-05-14 appendix.qmd
Updated Figure 1

```{r}
# Code to draw illustration for notation
start_pt <- all_centroids_df |> 
  filter(hexID == 1)
d_rect <- tibble(x1min = 0, 
                 x1max = 1,
                 x2min = 0,
                 x2max = diff(scurve_scaled_obj$nldr_obj$lim2)/diff(scurve_scaled_obj$nldr_obj$lim1))

@@

a1 <- tibble(x = all_centroids_df$c_x[4],
             xend = all_centroids_df$c_x[5],
             y = all_centroids_df$c_y[21],
             yend = all_centroids_df$c_y[21],
             label = expression(a[1]))
a2 <- tibble(x = all_centroids_df$c_x[27],
             xend = all_centroids_df$c_x[27],
             y = all_centroids_df$c_y[27],
             yend = all_centroids_df$c_y[33],
             label = expression(a[2]))
hex_param_vis <- ggplot() + 
    geom_polygon(data = hex_grid, 
                        aes(x = x, 
                            y = y, 
                            group = hex_poly_id),
                 fill = "white", 
                 color = "#bdbdbd") +
    geom_point(data = all_centroids_df, aes(
      x = c_x, 
      y = c_y), 
      color = "#31a354", size = 0.9) +
    geom_point(data = start_pt, aes(x = c_x, 
                                    y = c_y), 
               color = "black") + 
    geom_rect(data=d_rect, 
              aes(xmin = x1min - c_start$s1,# - c_start$s1, 
                  xmax = x1max - c_start$s1,# - c_start$s1, 
                  ymin = x2min - c_start$s2,# - c_start$s2, 
                  ymax = x2max - c_start$s2),# - c_start$s2), 
              fill = "white", 
              color = "black", 
              alpha = 0, 
              linewidth = 0.7) +
    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
                                y=x2min - c_start$s2)) + 
    geom_point(data=d_rect, aes(x=x1max - c_start$s1, 
                                y=x2min - c_start$s2)) + 
    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
                                y=x2max - c_start$s2)) + 
    annotate("text", x=d_rect$x1min - c_start$s1, 
                     y=d_rect$x2min - c_start$s2,
                     label = "(0,0)", 
             hjust=-0.1, vjust=-0.3) + 
    annotate("text", x=d_rect$x1max - c_start$s1, 
                     y=d_rect$x2min - c_start$s2,
                     label = "(0,1)", 
             hjust=1.1, vjust=-0.3) + 
    annotate("text", x=d_rect$x1min - c_start$s1, 
                     y=d_rect$x2max - c_start$s2,
                     label = expression(group("(", 
                        list(0, y[2][max]),")")), 
            hjust=-0.1, vjust=1.2) + 
    geom_segment(data=d_rect, aes(
      x = x1min  - c_start$s1, # 0 - 0.03, 
      y = -0.35, 
      xend = x1max - c_start$s1, #1 - 0.03, 
      yend = -0.35), #-0.35),
      arrow = arrow(length = unit(0.03, "npc"),
                               ends = "both"), 
                 color = "black")+
    annotate("text", x=0.5, y=-0.45, 
             label = expression(r[1]), color = "black") +
    geom_segment(data=d_rect, aes(
      x = -0.25, 
      y = x2min - c_start$s2, #0 - 0.05, 
      xend = -0.25, 
      yend = x2max - c_start$s2), #r2 - 0.05),
      arrow = arrow(length = unit(0.03, "npc"),
                       ends = "both"), 
                 color = "black")+ 
    annotate("text", x=-0.35, y=1, 
             label = expression(r[2]), color = "black") +
    geom_segment(data = a1, aes(
      x = x, #-0.1 + 0.2087578, 
      y = y, #-0.15, 
      xend = xend, #-0.1 + 0.2087578*2, 
      yend = yend), #-0.15),
      arrow = arrow(length = unit(0.03, "npc"),
        ends = "both"), 
        color = "black")+ # a1 = 0.2087578
    annotate("text", 
             x=(a1$x+a1$xend)/2, 
             y=a1$y, 
             label = expression(a[1]), 
             color = "black",
             vjust = 1.2) +
    geom_segment(data = a2, aes(
      x = x, #-0.15, 
      y = y, #-0.1*r2 + 0.1807896*2, 
      xend = xend, #-0.15, 
      yend = yend), #-0.1*r2 + 0.1807896*3),
      arrow = arrow(length = unit(0.03, "npc"),
                               ends = "both"), 
      color = "black") + # a2 = 0.1807896
    annotate("text", x=a2$x, y=(a2$y+a2$yend)/2, 
             label = expression(a[2]), 
             color = "black", hjust=-0.2) +
    annotate("text", x=-0.18, y=-0.25, 
      label = expression(group("(", list(s[1], s[2]), ")")),
      color = "black") +
  coord_equal()  

>>>>

```{r}
# Code to draw illustration for notation
start_pt <- all_centroids_df |> 
  filter(hexID == 1)

d_rect <- tibble(x1min = 0, 
                 x1max = 1,
                 x2min = 0,
                 x2max = diff(scurve_scaled_obj$lim2)/diff(scurve_scaled_obj$lim1))

@@

a1 <- tibble(x = all_centroids_df$c_x[4],
             xend = all_centroids_df$c_x[5],
             y = all_centroids_df$c_y[21],
             yend = all_centroids_df$c_y[21],
             label = expression(a[1]))
a2 <- tibble(x = all_centroids_df$c_x[27],
             xend = all_centroids_df$c_x[27],
             y = all_centroids_df$c_y[27],
             yend = all_centroids_df$c_y[33],
             label = expression(a[2]))

hex_param_vis <- ggplot() + 
    geom_polygon(data = hex_grid, 
                        aes(x = x, 
                            y = y, 
                            group = hex_poly_id),
                 fill = "white", 
                 color = "#bdbdbd") +
    geom_point(data = all_centroids_df, aes(
      x = c_x, 
      y = c_y), 
      color = "#31a354", size = 0.9) +
    geom_point(data = start_pt, aes(x = c_x, 
                                    y = c_y), 
               color = "black") + 
    geom_rect(data=d_rect, 
              aes(xmin = x1min - c_start$s1,# - c_start$s1, 
                  xmax = x1max - c_start$s1,# - c_start$s1, 
                  ymin = x2min - c_start$s2,# - c_start$s2, 
                  ymax = x2max - c_start$s2),# - c_start$s2), 
              fill = "white", 
              color = "black", 
              alpha = 0, 
              linewidth = 0.7) +
    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
                                y=x2min - c_start$s2)) + 
    geom_point(data=d_rect, aes(x=x1max - c_start$s1, 
                                y=x2min - c_start$s2)) + 
    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
                                y=x2max - c_start$s2)) + 
    annotate("text", x=d_rect$x1min - c_start$s1, 
                     y=d_rect$x2min - c_start$s2,
                     label = "(0,0)", 
             hjust=-0.1, vjust=-0.3, size = 8) + 
    annotate("text", x=d_rect$x1max - c_start$s1, 
                     y=d_rect$x2min - c_start$s2,
                     label = "(0,1)", 
             hjust=1.1, vjust=-0.3, size = 8) + 
    annotate("text", x=d_rect$x1min - c_start$s1, 
                     y=d_rect$x2max - c_start$s2,
                     label = expression(group("(", 
                        list(0, y[2][max]),")")), 
            hjust=-0.1, vjust=1.2, size = 8) + 
    geom_segment(data=d_rect, aes(
      x = x1min  - c_start$s1, # 0 - 0.03, 
      y = -0.35, 
      xend = x1max - c_start$s1, #1 - 0.03, 
      yend = -0.35), #-0.35),
      arrow = arrow(length = unit(0.03, "npc"),
                               ends = "both"), 
                 color = "black")+
    annotate("text", x=0.5, y=-0.45, 
             label = expression(r[1]), color = "black", size = 8) +
    geom_segment(data=d_rect, aes(
      x = -0.25, 
      y = x2min - c_start$s2, #0 - 0.05, 
      xend = -0.25, 
      yend = x2max - c_start$s2), #r2 - 0.05),
      arrow = arrow(length = unit(0.03, "npc"),
                       ends = "both"), 
                 color = "black")+ 
    annotate("text", x=-0.35, y=1, 
             label = expression(r[2]), color = "black", size = 8) +
    geom_segment(data = a1, aes(
      x = x, #-0.1 + 0.2087578, 
      y = y, #-0.15, 
      xend = xend, #-0.1 + 0.2087578*2, 
      yend = yend), #-0.15),
      arrow = arrow(length = unit(0.03, "npc"),
        ends = "both"), 
        color = "black")+ # a1 = 0.2087578
    annotate("text", 
             x=(a1$x+a1$xend)/2, 
             y=a1$y, 
             label = expression(a[1]), 
             color = "black",
             vjust = 1.2, size = 8) +
    geom_segment(data = a2, aes(
      x = x, #-0.15, 
      y = y, #-0.1*r2 + 0.1807896*2, 
      xend = xend, #-0.15, 
      yend = yend), #-0.1*r2 + 0.1807896*3),
      arrow = arrow(length = unit(0.03, "npc"),
                               ends = "both"), 
      color = "black") + # a2 = 0.1807896
    annotate("text", x=a2$x, y=(a2$y+a2$yend)/2, 
             label = expression(a[2]), 
             color = "black", hjust=-0.2, size = 8) +
    annotate("text", x=-0.18, y=-0.25, 
      label = expression(group("(", list(s[1], s[2]), ")")),
      color = "black", size = 8) +
  coord_equal()  
```
